# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

First, the date is read from file, and then split by dates.
```{r loadData}
activity <- read.csv2(unz('activity.zip', 'activity.csv'), sep=',')
activityByDate <- split(activity, activity$date)
```

## What is mean total number of steps taken per day?

Using a little utility function here, which takes one day as an argument.

```{r meanTotalStepsPerDay, fig.height=4}
stepsPerDay <- function(day) {
  return(sum(day$steps, na.rm=TRUE))
}
stepsPerDate <- sapply(activityByDate, stepsPerDay)

hist(stepsPerDate, main="Histogram of total steps per day", xlab="Steps per day")
```

The mean number of steps in the data is `r mean(stepsPerDate)` while the median `r median(stepsPerDate)`.

## What is the average daily activity pattern?

```{r avgDailyPattern, fig.height=3}
stepsPerInterval <- function(interval) {
  return(sum(interval$steps, na.rm=TRUE))
}

avgStepsPerInterval <- function(interval) {
  return(mean(interval$steps, na.rm=TRUE))
}

activityPerInterval <- split(activity, activity$interval)
plot(sapply(activityPerInterval, avgStepsPerInterval), type='l', main="Activity per interval", xlab="Interval number", ylab="Steps")
```

The most active interval on average is the interval number `r which.max(sapply(activityPerInterval, stepsPerInterval))`.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
