# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

First, the date is read from file, and then split by dates.
```{r loadData}
activity <- read.csv2(unz('activity.zip', 'activity.csv'), sep=',')
activityByDate <- split(activity, activity$date)
```

## What is mean total number of steps taken per day?

Using a little utility function here, which takes one day as an argument.

```{r meanTotalStepsPerDay, fig.height=4}
stepsPerDay <- function(day) {
  return(sum(day$steps, na.rm=TRUE))
}
stepsPerDate <- sapply(activityByDate, stepsPerDay)

hist(stepsPerDate, main="Histogram of total steps per day", xlab="Steps per day")
```

The mean number of steps in the data is `r mean(stepsPerDate)` while the median `r median(stepsPerDate)`.

## What is the average daily activity pattern?

```{r avgDailyPattern, fig.height=3}
stepsPerInterval <- function(interval) {
  return(sum(interval$steps, na.rm=TRUE))
}

avgStepsPerInterval <- function(interval) {
  return(mean(interval$steps, na.rm=TRUE))
}

activityPerInterval <- split(activity, activity$interval)
plot(sapply(activityPerInterval, avgStepsPerInterval), type='l', main="Activity per interval", xlab="Interval number", ylab="Steps")
# plot(by(activity$steps, activity$interval, function(d) {mean(d, na.rm=TRUE)}), type='l', main="Steps per interval", xlab="Intercal number", ylab="Steps per interval")
```

The most active interval on average is the interval number `r which.max(sapply(activityPerInterval, stepsPerInterval))`.

## Imputing missing values

The data has total `r sum(!complete.cases(activity))` NA's. A new data set will now be created. This solution was found online from [StackOverflow](http://stackoverflow.com/questions/9322773/how-to-replace-na-with-mean-by-subset-in-r-impute-with-plyr).

```{r handleNas, fig.height=4}
activityNoNas <- activity
for (i in which(sapply(activityNoNas, is.numeric))) {
    activityNoNas[is.na(activityNoNas[, i]), i] <- mean(activityNoNas[, i],  na.rm = TRUE)
}

activityByDateNoNas <- split(activityNoNas, activity$date)
stepsPerDateNoNas <- sapply(activityByDateNoNas, stepsPerDay)
hist(stepsPerDateNoNas, main="Histogram of total steps per day (NA's converted)", xlab="Steps per day")
```

The mean number of steps in the data now, with NA's replaced with means is `r mean(stepsPerDateNoNas)` while the median `r median(stepsPerDateNoNas)`. As can be witnessed from the historam here, replacing all unknown data points (NA) with mean values creates bias and is not such a great strategy for imputing. It moved low values around the mean. Something like interpolation, *k* nearest neighbours, or some random variables which have the same standard variance as the non-NA data would be a better idea.

It would be intersting to calculate how much error was introduced by thus method.

## Are there differences in activity patterns between weekdays and weekends?

```{r weekends}
# create a new column to indicate weekends (rude to manipulate date?)
activityNoNas$weekend <- weekdays(as.Date(activityNoNas$date)) %in% c('lauantai', 'sunnuntai')
# temporary data frame
df <- aggregate(activityNoNas$steps, list(activityNoNas$weekend, activity$interval), mean)
par(mfrow=c(2:1))
plot(df$x[df$Group.1 == FALSE], type='l', main="Weekdays", ylab="Activity", ylim=c(0, 250))
plot(df$x[df$Group.1 == TRUE], type='l', main="Weekend", ylab="Activity", ylim=c(0, 250))
```

The data shows on the weekend the person sleeps longer, and is on the move thoughtout the day, more evenly than during the week. Perhaps the peak during the week at around 10 is lunchtime?